An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-27 paper/1.introduction.tex
small fixes

In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
% A Survey to View Update Problem 
As a solution for this problem, bidirectional transformation (BX) is introduced. As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}. BX provides two functions: $get$ and $put$. Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.

@@

Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, he or she modified the view to \texttt{100} from \texttt{1}.
From this updated view, how can we obtain new updated source? This is ``the view update problem.'' For doing this, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
%\texttt{put pHead [1,2] 8 = [8,2]}

In BX programs, composition of programs is essential for writing various interesting evaluations. Because recursions are achieved by compositions, large number of programs include compositions.
In the BX lens papers \cite{} compositions are also known as central part of BX programs.
In this section we use a tiny BX program that contains compositions as a running example, because the programs including recursions can be complicated easily. The running examples are two versions of three composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).

@@

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section 2.

@@

\vspace{2mm}
In very-well behaved BX programs, 
$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
\vspace{2mm}

\begin{figure}[!t]
  \centering
  \includegraphics[height=5cm]{./fig/fig3.eps}
  \caption{Evaluating $phead \circ phead \circ phead$ by keeping intermediate states}
  \label{fig:eval-comp-phead-2}
\end{figure}

>>>>

%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time \cite{viewupdate}.
% A Survey to View Update Problem 
%As a solution for this problem, bidirectional transformation (BX) is introduced.
The synchronization of data is a common problem. In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{viewupdate}. Bidirectional transformation (BX) is an approach to solve this problem.
As an example, let us consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}. BX provides two functions: $get$ and $put$. Figure \ref{fig:eval-phead} denotes the behaviors of these functions when evaluating $phead$.

@@

Here, let \texttt{[1,2]} be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, the user
modified the view
from \texttt{1} to \texttt{100}.
From this updated view, how can the source be updated?
This is ``the view update problem.''
To solve this problem, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list \texttt{[100,2]} from the updated view (\texttt{v'}) and the original source (\texttt{s}).
Note that the update to s' uses s and v' as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

The composition of BX programs is a fundamental construct to build more complex BX programs. 
%In BX programs, composition of programs is essential for writing various interesting evaluations.
Because recursions are achieved by compositions, large number of programs include compositions.
In the BX lens papers \cite{} compositions are also known as central part of BX programs.
In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily. The running examples are two versions of three composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).

@@

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{josh_pepm} \cite{josh_popl}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section 2.

@@

\vspace{2mm}
In very-well behaved \cite{} BX programs, 
$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
\vspace{2mm}

\begin{figure}[!t]
  \centering
  \includegraphics[height=5cm]{./fig/fig3.eps}
  \caption{Evaluating $phead \ \circ \ phead \ \circ \ phead$ by keeping intermediate states}
  \label{fig:eval-comp-phead-2}
\end{figure}

### 2019-11-27 paper/2.minbigul.tex
small changes
small fixes

The \textsc{GetPut} [?] law means that if there is no change to the view, there should be no change to the source. The \textsc{PutGet} [?] law means that we can recover the modified view by applying the forward transformation to the updated source.

minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws. In addition, we restrict adaptive cases of BiGUL on minBiGUL. Then $put$ and $get$ satisfy one more law, \textsc{PutPut} [?], like the following:\\
    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$

The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates. Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved [?].

@@

We define the semantics of $put$ and $get$ as in definitions 1 and 2 respectively. Instead of using $v'$ in the $put$ direction like figures \ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view. The later definitions also use this convention.

>>>>

The \textsc{GetPut} \cite{putget} law means that if there is no change to the view, there should be no change to the source. The \textsc{PutGet} \cite{putget} law means that we can recover the modified view by applying the forward transformation to the updated source.

minBiGUL inherited from BiGUL also supports transformations $put$ and $get$ which are satisfy two above laws. In addition, we restrict adaptive cases of BiGUL on minBiGUL. Then $put$ and $get$ satisfy one more law, \textsc{PutPut} \cite{putput}, like the following:\\
    $\tab \putbx{bx}{(\putbx{bx}{s}{v'})}{v} = \putbx{bx}{s}{v} \qquad [\textsc{PutPut}]$

The \textsc{PutPut} law means that a source update should overwrite the effect of previous source updates. Due to the satisfaction of three laws, \textsc{GetPut}, \textsc{PutGet} and \textsc{PutPut}, minBiGUL is very well-behaved.

@@

We define the semantics of $put$ and $get$ as in definitions 1 and 2 respectively. Instead of using $v'$ in the $put$ direction like figures \ref{fig:eval-phead}, \ref{fig:eval-comp-phead} and \ref{fig:eval-comp-phead-2}, we just simply use notation $v$ as the updated view. The later definitions also use this convention. \textcolor{red}{$Leftarrow$ means assignments.}

### 2019-11-27 paper/6.related-work.tex
small fixes

Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we do not use this way.
  Instead of that, we introduce another way: using tupling \cite{}, lazy update and lazy computation \cite{}. Tupling \cite{tupling} is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation \cite{} is also a well-known and widely-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.

>>>>

Memoization \cite{} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is fusion \cite{}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
  Instead of that, we introduce another way that uses tupling \cite{tupling}, lazy update and lazy computation \cite{}. Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.
